Reducing Linearizability of Priority Queues and More Data Structures to State Reachability

### 2017-04-20 co-regular-EPQ.tex
first part of co-regular

\begin{restatable}{lemma}{priExecutionIsEnough}
\label{lemma:pri execution is enough}
Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ be a data-differentiated execution. Then, $e$ is not $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is not $\Gamma$-linearizable, for some maximal priority $p$ in $e$.
\end{restatable}
\begin {proof} (Sketch)
TODO UPDATE THE FOLLOWING
To prove the $\textit{only if}$ direction, given $e \sqsubseteq l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$. Since there is no restriction to $p$-incomparable operations in $u$, $v$ and $w$, we can safely erase all $p$-incomparable values. Similar case holds for $\Gamma=\mathsf{UnmatchedMaxPriority}$.

To prove the $\textit{if}$ direction, we use a lemma in Appendix, which said that if the projection of an execution into a set $O_1$ of operations has linearization $l_1$ and the projection of the execution into $O_2$, the remanning operations, has linearization $l_2$, then the execution has a linearization whose projection into $O_1$ is $l_1$ and projection into $O_2$ is $l_2$.

>>>>

\begin{restatable}{lemma}{priExecutionIsEnough}
\label{lemma:pri execution is enough}
Let $\Gamma\in \{\mathsf{UnmatchedMaxPriority}, \mathsf{MatchedMaxPriority}\}$ and $e$ be a data-differentiated execution. Then, $e$ is not $\Gamma$-linearizable if and only if $e\vert_{\preceq p}$ is not $\Gamma$-linearizable, for some maximal priority $p$ in $e$.
\end{restatable}
\begin {proof} (Sketch)
To prove the $\textit{only if}$ direction, let $e$ be an execution linearizable w.r.t. $l = u \cdot \textit{put}(x,p) \cdot v \cdot \textit{rm}(x) \cdot w \in \mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$. Since $\mathsf{MatchedMaxPriority}\mathsf{\text{-}Seq}(s,\alpha)$ imposes no restriction on the operations in $u$, $v$ and $w$ with priorities incomparable to $p$, erasing all these operations results in a sequential execution which still satisfies this property. Similarly, for $\Gamma=\mathsf{UnmatchedMaxPriority}$.

The $\textit{if}$ direction follows from the fact that if the projection of an execution to a set of operations $O_1$ has a linearization $l_1$ and the projection of the same execution to the remaining set of operations has a linearization $l_2$, then the execution has a linearization which is defined as an interleaving of $l_1$ and $l_2$ (see Appendix~\ref{} for more details).
